---
title:  "MSDS_6306_Case_Study_1"
author: "Jim Brittain"
date:   "October 26, 2016"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Load required packages (commented out because previously loaded)
#install.packages("ggplot2")
library(ggplot2)

## Adjust working directory as needed
MainDir <- "D:\\Users\\Jim\\SMU\\MSDS 6306 DOING DATA SCIENCE\\GIT\\MSDS_6306_Case_Study_1"
setwd(MainDir)
```
### Introduction
In this analysis report we plan to look at the possible relation of various countries’ Gross Domestic Product (GDP) and average income for the given country.  We will be using data available from the World Bank’s website: data.worldbank.org.  We plan to do some high-level analysis of these two variables for the various countries.   It is our hope that this analysis may show some promise for further statistical analysis in the future.

### Call each code section's make file

### Download the data files
```{r download files, echo=TRUE, tidy=TRUE}
source(file.path(MainDir,"Code","Download_Files.R"), echo=TRUE, keep.source=TRUE)
```

### Read and Clean files
```{r Read and Clean files, echo=TRUE, tidy=TRUE}
source(file.path(MainDir,"Code","ReadNClean_Files.R"), echo=TRUE, keep.source=TRUE)
```

### Merge and Analyze files
```{r Merge and Analize files, echo=TRUE}
source(file.path(MainDir,"Code","MergeNAnalyze_Files.R"), echo=TRUE, keep.source=TRUE)
```

## Results:

### Questions:

1. Merge the data based on the country shortcode. 
   How many of the IDs match?

Based on a left join we get a resulting data frame of 214 observations.  Using a intersection join to see how many of the ID variables matched we get 210 matched.


2. Sort the data frame in ascending order by GDP so United States is last. 
   What is the 13th country in the resulting data frame?
   
After sorting the data by GDP (ascending) we can see that St. Kitts and Nevis is the 13th observation.   

3. What are the average GDP rankings for the "High income: OECD" and 
   "High income: nonOECD" groups?
   
The mean (average) of the "High income: OECD" group is:   110.0667.
The mean (average) of the "High income: nonOECD" group is: 58.64865.     
   
   
4. Plot the GDP for all of the countries. 
   Use ggplot2 to color your plot by Income Group.

The plot illustrates that countries in the "High income: OECD" group also tend to have the highest GDP. It also shows a trend of lower incomes related to lower GDP.  Because the income categories are presented out of "true" order it is a little difficult to interpret.


5. Cut the GDP ranking into 5 separate quantile groups. 
   Make a table versus Income.Group.
   How many countries are Lower middle income but among the 38 nations with highest GDP?
   
The cross tabulation table shows a result of 9 countries where the income falls in the "Lower Middle" group and the income is in the 48 nations that fall into the top GDP quantile group.   
   
   
### Conclusion
Through this cursory analysis of various countries’ Gross Domestic Product (GDP) and average income for the given country we can glean some information of interest.  However, this is not a statistical analysis to determine if there is any kind of significant correlation of these two variables.  From our limited analysis it does appear to show some evidence that a country with a higher average income level; tends to have a higher GDP.  This could lead to more analysis to determine statistical significance of this correlation.  Because it is an observational study the statistical outcomes are limited due to unforeseen contributing variables.